---
title: "Class 8 Exercises"
output: html_document
date: "2024-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
```

## Exercises

### Easy

Do all the easy exercises from Chapter 9: **9E1** through **9E6**.

9E3:

The HMC does not work very well when the priors are weak and the data ditto (uninformative). The way I think about this is, that the 'particles' in the simulation will cover a very large flat area before turning around. They will not learn the shape of the posterior very effectively (but also, the posterior will have a very uninformative shape anyway, regardless of algorithm efficiency). 

9E4: 

The effective number of samples *n_eff* is the number of samples with optimal (that is with no or even anti) auto-correlation that would be necessary to obtain similar information as the actual number of samples did. Optimal samples gain maximum information for each step in the chain.

9E5:

1 from above.The r-hat is a diagnostic that can be used to assess how well the chains mixed. If the r-hat is high, this means that the chains did not mean very well. 

### Medium

#### 9M1

```{r}
# loading data
library(rethinking)
data(rugged)
d <- rugged
d$log_gdp <- log(d$rgdppc_2000)
dd <- d[ complete.cases(d$rgdppc_2000) , ]
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )
```

```{r}
# transforming variables
dat_slim <- list(
    log_gdp_std = dd$log_gdp_std,
    rugged_std = dd$rugged_std,
    cid = as.integer( dd$cid )
)
str(dat_slim)
```

```{r}
# estimating the posterior
m9.1 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) , data=dat_slim )

precis(m9.1)
```

```{r}
# changing the prior on sigma to a uniform distribution
m9.2 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dunif( 0 , 10 )
    ) , data=dat_slim )
```

```{r}
# changing the prior on sigma to an exponential distribution
m9.3 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) , data=dat_slim )
```

```{r}
dexp( 1 )

dexp(1)
```


```{r}
precis( m9.2 )
precis( m9.3 )
```

The changed priors doesn't change anything.

#### 9M2

```{r}

# Your solution herre

```

### Hard

#### 9H1

```{r}

# Your solution here

```

#### 9H2

```{r}

# Your solution herre

```
